---
title: "desc_var Vignette"
author: "Sahardid Mohamed"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{desc_var Vignette}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction

This vignette demonstrates how to use the `desc_var` function from the `descusmr` package.

# Usage

To get started, first load the `descusmr` package:

```{r}
library(develloppementfonction)
```


The dataset to be analyzed is passed through the `desc_var` function.
- table_title: Title of the descriptive table. Here, it is "test".
- var_group: The variable used for grouping the data. Here, it is "traitement".
- group_title: Title for the grouping variable column. Here, it is "traitement".
- group: Une description groupée et degroupée à la fois est réalisé avec l'option "ALL"
- DM: We sort the frequencies of the modalities in descending order and remove modalities with zero frequencies with the option DM = "tout". 


```{r}
data %>% desc_var(table_title = "test",
             var_group = "traitement",
             group_title = "traitement",
             group = "ALL", 
             DM = "tout")
```



